---
title: "Regime Types in Latam since 1945"
author: "Ana Luisa Bodevan"
format: html
editor: visual
---

## Objective

How did regimes changed and evolved in Latin America since 1945? This graph uses Cheihub and Vreeland (2010) classification of regimes according to various metrics explained in their paper "Democracy and Dictatorship Revisited" (2010).

This visualization is inspired and draws deeply from the incredible tutorial at [R for Political Science](https://rforpoliticalscience.com/2022/03/18/download-democracy-data-with-democracydata-pacakge-in-r/)

```{r}
#| echo: false

getwd()
setwd("C:/Users/anabo/OneDrive/Documentos/R/ggplot2")
getwd()
```

#### Packages

```{r}

#| label: setup

library(pacman)
pacman :: p_load('democracyData', 'tidyverse', 'magrittr', 'ggstream', 'ggthemes', 'MetBrewer')
```

#### Data

Dataset is from Cheihub and Vreeland (2010) aforementioned paper, dowloaded with the democracyData package by [Xavier Marquez](https://github.com/xmarquez)

```{r}

#| label: data wrangling 

pacl <- redownload_pacl()

pacl %<>% 
  mutate(regime_name = ifelse(regime == 0, "Parliamentary democracies",
       ifelse(regime == 1, "Mixed democracies",
       ifelse(regime == 2, "Presidential democracies",
       ifelse(regime == 3, "Civilian autocracies",
       ifelse(regime == 4, "Military dictatorships",
       ifelse(regime ==  5,"Royal dictatorships", regime))))))) %>%
  mutate(regime = as.factor(regime)) # create factor variables according to the codebook

americas <- subset(pacl, un_continent_name == "Americas") # create a new dataset for Latam

regime_types <- c("Parliamentary democracies", "Presidential democracies", 
                  "Mixed democracies", "Royal dictatorships", 
                  "Civilian autocracies", "Military dictatorships")

types <- americas %>%
  group_by(year, regime_name) %>%
  count() %>%
  filter(regime_name %in% regime_types) %>%
  arrange(year, desc(n)) %>%
  ungroup() # creates a dataset storing each year and the types of regimes 
```

#### Visualization

```{r}

#| label: plotting

palette <- met.brewer('Peru2', n = 6, type = 'discrete', direction = 1)

ggplot(types, aes(x = year, y = n, fill = regime_name)) +
  geom_area(position = "stack", color = "white") +
  scale_fill_manual(values = palette) +  # Use MetBrewer color palette
  labs(x = "Decade", y = NULL, fill = "Regime Type", title = "Regime Types in Latin America since 1945", subtitle = "by Ana Luisa Bodevan, inspired by R for Political Science\nSource: Cheihub and Vreeland (2010)") +  # Labels for axes and legend
  theme_minimal() +
  theme(panel.grid.major = element_blank(),  
        panel.grid.minor = element_blank(),
        legend.position = "bottom")

```
